Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-09-20 ba_resubmit/response_to_reviewers.tex

	\section*{Reviewer 3}
	\setcounter{responsectr}{0}

	The paper presents a computational variant of the method introduced by Sadinle (2017). Simulation results are presented and the variant is applied on a new, significantly larger data set. A significant part of the paper is a reproduction of the paper of Sadinle. The variant is introduced in section 4. It is simple but consequential. The simulations and applications show how the variant accelerates the computation relative to the original method and reveals the trade-offs.

	The value of the paper resides in that computational complexity continues to be a difficult obstacle in the application of Bayesian statistics. The proposals of the paper enable the application of Sadinle’s approach to an extent not possible before. It is important to present the research for what it is: An extension of the method of Sadinle motivated by computational considerations. Here Sadinle supersedes Fellegi-Sunter in the sense while the starting point of Sadinle’s research is Fellegi-sunter, the starting point of this paper is Sadinle’s work. This should be made clearer, beginning in the abstract and reiterated in the conclusion.

	\reply{%
	\textbf{Thank you for your positive feedback regarding our paper. We have clarified that we are motivation by the work of Sadinle and computational considerations. We have tried to make this more clear in both the abstract, introduction, and conclusion.}} \todo

### 2022-10-01 ba_resubmit/kundinger_resubmit.tex

\begin{abstract}
Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.  \textcolor{blue}{We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations. 
Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}
Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.

\textcolor{red}{BEKA/JERRY: Reviewer 2 wanted a specific reference to Sadinle in the abstract. I remember we actively decided not to do that. What do you think?} \textcolor{blue}{I added this back in. He only has one concern, so let's make sure to satisfy the person.}
\end{abstract}

@@

In many data analysis tasks, analysts seek to identify duplicate records across two databases. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}. In this article, we consider bipartite record linkage, which merges two databases that contain duplications across but not within the respective databases.

@@

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice. In many situations, we know that there are no duplications within a database, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost. \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on databases with more than 100 records. \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weight the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review. \texttt{BRL} was shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.

\textcolor{blue}{In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2),$ where $n_1$ and $n_2$ are the sizes of each database. The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}. In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.}

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs and (2) compute the pairwise record comparisons over large datasets via parallel computing. \textcolor{blue}{Moreever, we propose storage efficient indexing (SEI), which reduces memory costs.}
 \textcolor{blue}{More specifically, we use both hashing and SEI to increase the scale of the linkage tasks without the use of blocking. This can be useful when there is no reliable blocking field available or one desires estimates of model parameters for the entire sample in question. In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking. We illustrate our methodology on two case studies and two simulation scenarios, where Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.}

@@

In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section \ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and shows the loss function used to calculate the Bayes estimate for the bipartite matching. Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}. Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.

@@

\textcolor{red}{Consider two databases $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts.} In the context of bipartite matching, we assume that there are duplications across, but not within, each database. Under this framework, the set of matches across databases can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'', and record pairs that are estimated as non-matches as ``non-links''.

@@

The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables. For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absolute difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then can set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choice of $I_{fl}$ are application specific, which we discuss in our simulation and case studies.

%\textcolor{red}{(I don't love this right here, but I don't really know where else to put it. But if it won't cause a problem with the reviewers, I'll just leave it.)}
%\textcolor{blue}{Perhaps any parts that aren't needed can be removed. I'll take a stab at it.}

%In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}. In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
%
%In this paper, we use hashing and a new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking. This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question. In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking.

@@

Under the Fellegi Sunter model, we assume first, that comparison vectors are independent given their matching status, and second, that the match status of the record pairs are independent. This allows us to model the comparison data using mixture models in the following way:
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $p$ is the marginal probability that a record pair is a match. These $\bm{m}$ and $\bm{u}$ parameters are then used to calculate the likelihood ratio  
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
\end{align}
This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.

@@

To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model. He proposed the  beta distribution for bipartite matching, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the number of records in $\bm{X}_2$ that have a match in $\bm{X}_1$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can be slow when working on large linkage tasks.

@@

Recall the compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a match and $Z_j = n_1 + j$ indicates a non-match. In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper and flexible prior. We denote the fast beta prior as follows:
$$Z_j | \pi =
\begin{cases} 
	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
	1-\pi, &  z_j  = n_1 + j; \\
\end{cases}$$
$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}),$$
$\alpha_{\pi}$ and $\beta_{\pi}$ are fixed hyperparameters that encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}.$ Note the prior from \cite{wortman2019} can be viewed as a special case of the fabl prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.

\begin{remark}
Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match. One non-informative choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\). 
\end{remark}

\textcolor{blue}{Furthermore, note that it is common to encounter missing information in datasets used for record linkage. Whether information is missing in record $i \in \bf{X}_1$ or $j \in \bf{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We handle this issue by assuming that this missingness occurs at random (MAR, by \cite{LittleRubin2002}). With this assumption, we can marginalize over the missing data, and do all computation simply using the observed data. To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.}
\textcolor{red}{(I don't love this right here, but I don't really know where else to put it.)}

@@

@@

%The model can then be specified as follows:
With the \texttt{fabl} prior defined and our missingness assumptions addressed, we present the full model. Let $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|Z_j = i)$ for all fields $f$ and agreement levels $l$. Let $\bm{u}$ parameters be defined analogously, with $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$. Then, using the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$, our model is:
\begin{align*}
	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \\
	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}), \\
	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}), \\
	Z_j | \pi &=
	\begin{cases} 
		\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
		1-\pi, &  z_j  = n_1 + j; \\
	\end{cases} \\
	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}),
\end{align*}
%where we have used the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$ to denote the set of parameters governing the probabilities of agreement levels for matching and non-matching record pairs.

\textcolor{red}{I actually don't like the $\Phi$ notation. It seems unnecessary. The only reason it's here is because Sadinle used it. What do you guys think?}
\textcolor{blue}{I would leave it since they didn't complain unless you have a better idea.}

>>>>

\begin{abstract}
Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. \textcolor{red}{To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.} Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy \textcolor{red}{when compared to \texttt{BRL}.}
%
%\textcolor{blue}{We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.  Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}

\end{abstract}

@@

In many data analysis tasks, analysts seek to identify duplicate records across two datafiles. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}. In this article, we consider bipartite record linkage, which merges two datafiles that contain duplications across, but not within, the respective datafiles.

@@

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice. In many situations, we know that there are no duplications within a datafile, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost. \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on datafiles with more than 100 records. \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review. \texttt{BRL} was shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. \textcolor{red}{These contributions allow \texttt{fabl} to perform record linkage on much larger datafiles than previous Bayesian Fellegi Sunter models at significantly increased speed with minimal loss of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each datafile, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller datafile.}

%\textcolor{blue}{Moreever, we propose storage efficient indexing (SEI), which reduces memory costs.}
%\textcolor{blue}{More specifically, we use both hashing and SEI to increase the scale of the linkage tasks without the use of blocking. This can be useful when there is no reliable blocking field available or one desires estimates of model parameters for the entire sample in question. In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking. We illustrate our methodology on two case studies and two simulation scenarios, where Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.}

@@

In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and shows the loss function used to calculate the Bayes estimate for the bipartite matching. Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}. Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.

@@

Consider two datafiles $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts. In the context of bipartite matching, we assume that there are duplications across, but not within, each datafile. Under this framework, the set of matches across datafiles can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in datafile $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in datafile}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''

@@

The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables. For numerical data, we can use absolute difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We can then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, which we discuss in our simulation and case studies.

%For more complex measurements, we can take into account partial agreement to more richly characterize the comparison;

%\textcolor{red}{(I don't love this right here, but I don't really know where else to put it. But if it won't cause a problem with the reviewers, I'll just leave it.)}
%\textcolor{blue}{Perhaps any parts that aren't needed can be removed. I'll take a stab at it.}
%In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}. In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
%
%In this paper, we use hashing and a new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking. This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question. In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking.

@@

The Fellegi Sunter model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. This allows us to model the comparison data using mixture models in the following way:
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $p$ is the marginal probability that a record pair is a match. The $\bm{m}$ and $\bm{u}$ parameters are used to calculate the likelihood ratio  
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
\end{align}
This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.

%These unknown parameters are often estimated using the EM algorithm \cite{winkler_state_1999}.

@@

To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a prior distribution over the matching space. He proposed the  beta distribution for bipartite matching, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the number of records in $\bm{X}_2$ that have a match in $\bm{X}_1$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.

@@

In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper prior. We denote the fast beta prior as follows:
$$Z_j | \pi =
\begin{cases} 
	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
	1-\pi, &  z_j  = n_1 + j; \\
\end{cases}$$
$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).$$
This set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match. The parameter $\pi$ is given hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ that encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}.$

% I don't really like using remarks. If its not relevante enough for the main body, we can just remove. 
%\begin{remark}
%One non-informative choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\). Note the prior from \cite{wortman2019} can be viewed as a special case of the fabl prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable. 
%\end{remark}

@@

Lastly, we note that it is common to encounter missing information in datasets used for record linkage. Whether information is missing in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We handle this issue by assuming that this missingness occurs at random (MAR, by \cite{LittleRubin2002}). With this assumption, we can marginalize over the missing data, and do all computation simply using the observed data. To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.

@@

%The model can then be specified as follows:

\begin{table}[t!]
	\centering
	\begin{tabular}[t!]{ll}
		Symbol & Description \\
		\hline
		$\bm{X}_1, \bm{X}_2$ & datasets\\
		$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
		$j \in 1, \ldots, n_2 $ & index over records in $\bm{X}_2$\\
		$f \in 1, \ldots F$ & index over fields used for comparisons \\
		$l \in 1, \ldots L_f$ & index over agreement levels for field $f$ \\
		$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
		$\bm{\gamma}_{ij}$ & comparison vector for records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ \\
		$Z_j = i$ & records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ match \\
		$Z_j = n_1 + j$ & record $j \in \bm{X}_2$ has no match in $\bm{X}_1$ \\
		$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
		$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
		$\pi$ & probability that a record in $\bm{X}_2$ has a match in $\bm{X}_1$ \\
		\hline
	\end{tabular}\caption{Summary of model notation}\label{table_notation_1}
\end{table}
